Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV theory

### 2019-10-15 paper2_JNLS/paper.tex
NM

\title{Anomalous waves triggered by abrupt depth changes: experiments and truncated KdV theory}

@@

The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019} and Majda et al.~\cite{majda2019, majdaqi2019}. Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}. 
%We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. 
Novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory. Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}. In addition, we provide a more thorough treatment of the link between experimental and model parameters to facilitate comparison between the two. In particular, we flesh out details of non-dimensionalization, which were only briefly addressed in prior work for the sake of brevity.

@@

% Figure: Experimental Diagram and Histograms
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.80 \linewidth]{Figs/ExpDiagStats.pdf}
\caption{
(a) Experimental schematic. 
(b)--(c) Surface displacement measured at a representative upstream and downstream locations. (d)--(e) Corresponding histograms. Figure adapted from \cite{bolles2019}.}
\label{ExpDiagStats}
\end{center}
\end{figure}
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Data from column 5 in MasterTimeSeries, Delta theta = 1.38 degrees, which is in between runs 5 and 6.

@@

% Figure: Experimental Spatial Statistics
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.80 \linewidth]{Figs/ExpSpatialStats.pdf}
\caption{Wave statistics as they vary in space for several experiments of different driving amplitudes (see legend). (a)--(c) Standard deviation, skewness, and kurtosis of the surface displacement $\eta$. (d)--(f) The same for surface slope $\eta_x$. (a) $\etastd$, which sets a scale for wave amplitude, does not vary significantly crossing the ADC. (b)--(c) Skewness and kurtosis of $\eta$, however, show a strong response to the ADC. The surface slope measurements are negatively skewed and also show a large kurtosis downstream of the ADC. }
\label{ExpSpatialStats}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

Recall that $\Ham_{\Lambda}$ is composed of the cubic and quadratic components $\Hthree$ and $\Htwo$. Due to the odd symmetry of $\Hthree$, it is easy to see that $\meanz{\Hthree} = 0$. The quadratic component, however, requires closer inspection. Due to Parseval's identity, $\Htwo$ is given by
\begin{equation}
\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx = 2 \pi \sum_{k=1}^{\Lambda} k^2 \abs{\uhat_k}^2
\end{equation}
Due to the constraint $\En[\uL] = 1$, the equipartioned microstate is given by 
\begin{align}
&\abs{\uhat_k}^2 \approx \frac{1}{2 \pi \Lambda}	 \qquad \mbox{\em for equipartition of energy}
\end{align}
Then the expected value of $\Htwo$ under the uniform measure is
\begin{equation}
\label{HtwoExpect}
\mean{\Htwo}_0 = 2 \pi \sum_{k=1}^{\Lambda} \abs{k}^ 2 \mean{\abs{\uhat_k}^2}_0 \sim \frac{1}{3} \Lambda^2
\end{equation}
where we have used the identity for the Gauss-like sum
\begin{equation}
\sum_{k=1}^{n} \abs{k}^ 2 = \frac{1}{6} n(n+1)(2n+1) \approx n^3/3
\end{equation}

	Importantly, \eqref{HtwoExpect} shows that the expected value of $\Htwo$ grows like $\Lambda^2$, which appears to be problematic for obtaining independence as $\Lambda \to \infty$. However, $\Htwo$ enters $\Ham$ in product with the coefficient $C_2$, which itself scales as $C_2 \sim \lamfac^{-2}$. Hence, obtaining the desired asymptotic independence with respect to $\Lambda$ simply requires that $\lamfac$ grow proportionally to $\Lambda$, as was already argued on physical grounds in Section \ref{tKdVSec}. 
Thus, any of the choices $\lamfac = \Lambda$, $\Lambda/2$, or $\Lambda/4$, discussed in that section would be valid. In particular, if $\lamfac = \Lambda$, then the characteristic wavelength in experiments corresponds to the smallest resolved wavelength in the dynamical system. It is perhaps more sensible to choose $\lamfac$ to be an intermediate value between 1 and $\Lambda$, so that some scales both larger than and smaller than the characteristic wavelength $\lam$ are resolved. As default, we will choose $\lamfac = \Lambda/2$, so that on a log-scale, the characteristic wavelength lies directly in the middle of the resolved wavelengths.

@@

	Next, we aim to make the same comparison for free-surface slopes. We note that the derivative, $\partial \eta / \partial x$, is already a dimensionless quantity with a simple physical interpretation, namely the slope of the free surface, whereas the interpretation of $\partial u/\partial \tilde{x}$ is tied to the characteristic values $\ampscale$ and $\lengthscale$.  
We therefore convert the theoretical calculated slopes values back to physical slopes via
\begin{equation}
\pd{\eta}{x} = 2 \pi^{3/2} \frac{\etastd}{\lamupdn} \pd{\uupdn}{\tilde{x}}
%\frac{\ampscale \lamfac}{\lengthscale} \pd{\uupdn}{\tilde{x}}
\end{equation}
So that the peak experimental wavelength $\lam$ (corresponding to $\freqp = 2$ Hz) would be mapped to the dominating TKdV spectral mode of $k = 1$,  this conversion formula has not been corrected with the extra factor of $\lamfac$ present in \eqref{scales}.

@@

	Figure \ref{slopehist} shows the distributions of surface slope that result from experiments (top) and theory (bottom). These histograms show intriguing similarities and differences. First, the upstream slope distribution is captured well by the TKdV simulations, both in its nearly symmetric shape and its scale. We note that, while the standard deviation of displacement was input into the TKdV theory, the slope standard deviation was not. Downstream, both the experimental and theoretical slope distributions  spread significantly. Interestingly, in the experiments, it is not the standard deviation that increases significantly upon crossing the depth change ($\std(\eta_x)$ increases from 0.15 to 0.17), but instead the excess kurtosis, which jumps from a value of 0.46 upstream to 3.4 downstream. Likewise, in the TKdV theory, $\std(\eta_x)$ grows from 0.13 upstream to 0.33 downstream, and $\kurt(\eta_x)$ grows from -0.04 to 0.50. Thus, the theoretically predicted jump in kurtosis, while not as extreme as that measured in experiments, is quite significant. These elevated levels of kurtosis are associated with the distinct appearance of the distribution, having flat tails and somewhat sharp centers.

	Distinct differences in the shapes of the downstream slope distributions, however, are visible. While the theoretical distributions remain symmetric, the experimental distributions skew towards negative slope. As mentioned earlier, this skewness is consistent with a right-moving wave of steep leading surface (i.e.~negative slope), which is a phenomenon apparently not captured by the TKdV simulations.
Finally, it should be commented that the experimental measurements of slope involve numerical differentiation of the free surface extracted from optical images, which naturally introduces error. Furthermore, high-order statistics, such as kurtosis, can be very sensitive to noise and measurement error. Hence, the experimental statistics of slope, while highly valuable, should be regarded with some level of caution.

%----------------------------------------------------------------------------%
% Wave dynamics
%----------------------------------------------------------------------------%
\subsection{Wave dynamics}

@@

These observations can be rationalized with some simple analysis of the structure of TKdV. Upstream of the ADC, since $\drat = 1$, the nonlinear effects are relatively weak and the dynamics are dominated by dispersion. Linearizing Eq.~\eqref{TKdV} and introducing the ansatz $u_k = e^{i(kx - \omega_k t)}$ produces the dispersion relation for angular frequency $\omega_k = -C_2 k^3$, or for phase velocity $c_k = \omega_k/k = -C_2 k^2$ \cite{majdaqi2019}. In particular, the phase velocity is strictly negative, implying that waves can only propagate leftwards, as is consistent with Fig.~\ref{trajectories}(a).
% Further details, see page 6 of Majda Qi, J Stat Phys.
%$C_3 = 0.20$ and $C_2 = 0.024$.

@@

	We remind the reader that the KdV framework tracks the long-time evolution of waves in a reference frame moving with the characteristic speed $c = \sqrt{g \depth}$ from linear theory. Hence, the above discussion of left-or-right going waves cannot be interpreted in the context of the experiments, which take a fixed reference frame, without an appropriate Galilean transformation. Hence, in the laboratory frame, all waves indeed propagate unidirectionally, from left to right, with a speed that is near $c$. The directions and speeds calculated by the TKdV framework simply determines how much the true wave speed deviates from $c$.

@@

	These estimates raise an important question: if a distance of 22 meters is required for significant wave evolution under the KdV framework, how do the experiments exhibit substantial changes in wave statistics a much shorter distance downstream of the ADC? The results shown in Fig.~\ref{skew-kurt} help resolve this question. This figure shows the evolution of surface-displacement skewness (left) and kurtosis (right) computed by the same TKdV deterministic simulations as pictured in Fig.~\ref{trajectories}. The bold black curve shows the ensemble mean (ensemble size of 1000), while the faint gray curves show the skewness and kurtosis of 100 individual trajectories to give a sense for the variation involved. Importantly, the skewness and kurtosis evolve on a much shorter timescale than the aforementioned $t=7$. More precisely, skewness and kurtosis have already saturated to their asymptotic value by $t = 1$, and reach half of that value by $t = 0.18$. These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order as the relevant distances in the experiments. Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter its statistical distributions, and this reconfiguration occurs on a much shorter timescale than that required for waves to cross the entire periodic domain.

 %----------------------------------------------------------------------------%
% Skewness Ratio 
%----------------------------------------------------------------------------%
\subsection{Predicted outgoing skewness and experimental confirmation}

This section discusses arguably the most novel single result of the manuscript. In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}. More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$. Below, we recap the derivation of this formula and then test the prediction against new, direct experimental measurements.

%\subsection{Explicit formula for outgoing skewness}

@@

	Our next task is to convert this formula to physical variables so that it can be tested against experimental data. First, following the definition of $\Hthree$ in Eq.~\eqref{H3H2} and the definition, $u = \eta/(\pi^{1/2} \etastd)$, a straightforward calculation gives
\begin{align}
\label{H3convert}
\meandn{\Hthree} = \frac{\pi}{3} \meandn{u^3} = 
\frac{1}{3 \pi^{1/2}} \frac{\meandn{\eta^3}}{ \etastd^3} = 
\frac{1}{3 \pi^{1/2}} \skwdn(\eta)
\end{align}
This equation expresses a direct relationship between the expected value of $\Hthree$ and the outgoing skewness of surface displacement. Next, we must convert the $\Htwo$-quantities, which involve surface slope. 
Using the definition of $\Htwo$ in Eq.~\eqref{H3H2} and $\lengthscale_+$ in Eq.~\eqref{scales} gives
\begin{align}
\label{H2convert}
&\meanupdn{\Htwo} = \pi \meanupdn{ u_x^2} = 
\left( \frac{\lamfac \lamdn}{2 \pi \etastd} \right)^2 \varupdn(\eta_x)
\end{align}
We note that the downstream scale $\lengthscale_+$ is used in this conversion, since it is the downstream Hamiltonian that is matched in \eqref{statmatch}. This formula links the expected value of $\Htwo$ to the variance of surface slope.

@@

	Fortunately, it is a relationship that can be tested directly against the experimental measurements, specifically the type of data that was presented in Fig.~\ref{ExpSpatialStats}. The the prediction draws focus to the spatial variation of displacement skewness, which we replot for convenience in Fig.~\ref{SkewRat}(a), and the variance of surface slope, shown in Fig.~\ref{SkewRat}(b). To test \eqref{SkewPred}, we must select a representative downstream position to evaluate $\skwdn(\eta)$ and $\vardn(\eta_x)$, and a representative upstream position for $\varup(\eta_x)$. In short, we select the same representative locations as were used to evaluate the histograms in Figs.~\ref{ExpDiagStats}, \ref{uhist}, and \ref{slopehist}, namely $x = 15$ cm downstream and $x = -9$ cm upstream. These positions are indicated by the blue and green vertical dashed lines in Fig.~\ref{SkewRat}(a)--(b). Recall that $x = 15$ cm corresponds to the peak of the downstream skewness and kurtosis. Meanwhile, we see in Fig.~\ref{SkewRat}(b) that $x = -9$ cm corresponds to a slight dip in $\var(\eta_x)$.

	With these locations selected, we evaluate the ratio $\skw(\eta) / \Delta \var(\eta_x)$ for each experiment and plot the result against dimensionless wave amplitude, $\epsup = \etastd / \dup$ on a log-log scale in Fig.~\ref{SkewRat}(c). While the data shows significant variation for small amplitudes, it shows a well defined decreasing trend at larger amplitudes. In particular, the red dashed line shows the power-law $\epsup^{-3}$ predicted by \eqref{SkewPred}, which captures the experimental trend remarkably well.

@@

>>>>

\title{Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV theory}

@@

The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019} and Majda et al.~\cite{majda2019, majdaqi2019}. Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}. 
We first provide a more thorough treatment of the link between experimental and model parameters in order to facilitate a detailed comparison between the two. Next, we present new experimental data, most notably on {\em surface slope} statistics, which motivates parallel examination of the theory.  Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory. Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}.

% OLD VERSION
\begin{comment}
We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. 
Novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory. Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}. 
In addition, we provide a more thorough treatment of the link between experimental and model parameters in order to facilitate a detailed comparison between the two. 
In particular, we flesh out details of non-dimensionalization, which were only briefly addressed in prior work for the sake of brevity.
\end{comment}

@@

% Figure: Experimental Diagram and Histograms
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.80 \linewidth]{Figs/ExpDiagStats.pdf}
\caption{
(a) Experimental schematic: randomized waves are generated by a pivoting paddle and propagate over a step in bottom topography.
(b)--(c) Surface displacement measured at representative locations upstream and downstream of the abrupt depth change (ADC). (d)--(e) Corresponding histograms showing symmetric upstream statistics and highly skewed downstream statistics. Figure adapted from \cite{bolles2019}.}
\label{ExpDiagStats}
\end{center}
\end{figure}
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Data from column 5 in MasterTimeSeries, Delta theta = 1.38 degrees, which is in between runs 5 and 6.

@@

% Figure: Experimental Spatial Statistics
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.80 \linewidth]{Figs/ExpSpatialStats.pdf}
\caption{Wave statistics as they vary in space for several experiments of different driving amplitudes (see legend). (a)--(c) Standard deviation, skewness, and kurtosis of the surface displacement $\eta$. (d)--(f) The same for surface slope $\eta_x$. (a) $\etastd$, which sets a scale for wave amplitude, does not vary significantly crossing the ADC. (b)--(c) Skewness and kurtosis of $\eta$, however, show a strong response to the ADC. (d)--(f) The measurements of surface slope, $\eta_x$, are negatively skewed and also exhibit large kurtosis downstream of the ADC. }
\label{ExpSpatialStats}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

Recall that $\Ham_{\Lambda}$ is composed of the cubic and quadratic components $\Hthree$ and $\Htwo$. Due to the odd symmetry of $\Hthree$, it is easy to see that $\meanz{\Hthree} = 0$. The quadratic component, however, requires closer inspection. Due to Parseval's identity, $\Htwo$ can be written as
\begin{equation}
\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx = 2 \pi \sum_{k=1}^{\Lambda} k^2 \abs{\uhat_k}^2
\end{equation}
Due to the constraint $\En[\uL] = 1$, the equipartioned microstate is given by 
\begin{align}
&\abs{\uhat_k}^2 \approx \frac{1}{2 \pi \Lambda}	 \qquad \mbox{\em for equipartition of energy}
\end{align}
Then the expected value of $\Htwo$ under the uniform measure is
\begin{equation}
\label{HtwoExpect}
\mean{\Htwo}_0 = 2 \pi \sum_{k=1}^{\Lambda} {k}^ 2 \mean{\abs{\uhat_k}^2}_0 \sim \frac{1}{3} \Lambda^2
\end{equation}
where we have used the identity for the Gauss-like sum
\begin{equation}
\sum_{k=1}^{n} {k}^ 2 = \frac{1}{6} n(n+1)(2n+1) \approx \frac{1}{3} n^3
\end{equation}

	Importantly, \eqref{HtwoExpect} shows that the expected value of $\Htwo$ grows like $\Lambda^2$, which appears problematic for obtaining independence as $\Lambda \to \infty$. However, $\Htwo$ enters $\Ham$ in product with the coefficient $C_2$, which itself scales as $C_2 \sim \lamfac^{-2}$. Hence, obtaining the desired asymptotic independence with respect to $\Lambda$ simply requires that $\lamfac$ grow proportionally to $\Lambda$, as was already argued on physical grounds in Section \ref{tKdVSec}. 
Thus, any of the choices $\lamfac = \Lambda$, $\Lambda/2$, or $\Lambda/4$, discussed in that section would be valid. In particular, if $\lamfac = \Lambda$, then the characteristic wavelength in experiments corresponds to the smallest resolved wavelength in the dynamical system. It is perhaps more sensible to choose $\lamfac$ to be an intermediate value between 1 and $\Lambda$, so that some scales both larger than and smaller than the characteristic wavelength $\lam$ are resolved. As default, we will choose $\lamfac = \Lambda/2$, so that on a log-scale, the characteristic wavelength lies directly in the middle of the resolved wavelengths.

@@

	Next, we aim to make the same comparison for free-surface slopes. We note that the derivative, $\partial \eta / \partial x$, is already a dimensionless quantity with a simple physical interpretation, namely the slope of the free surface, whereas the interpretation of $\partial u/\partial \tilde{x}$ is tied to the characteristic values $\ampscale$ and $\lengthscale$.  
We therefore convert the theoretical calculated slopes values back to physical slopes via
\begin{equation}
\pd{\eta}{x} = 2 \pi^{3/2} \frac{\etastd}{\lamupdn} \pd{\uupdn}{\tilde{x}}
%\frac{\ampscale \lamfac}{\lengthscale} \pd{\uupdn}{\tilde{x}}
\end{equation}
So that the peak experimental wavelength $\lam$ (corresponding to $\freqp = 2$ Hz) would be mapped to the dominating TKdV spectral mode, $k = 1$,  this conversion formula has not been corrected with the extra factor of $\lamfac$ present in \eqref{scales}.

@@

	Figure \ref{slopehist} shows the distributions of surface slope that result from experiments (top) and theory (bottom). These histograms show intriguing similarities and differences. First, the upstream slope distribution is captured well by the TKdV simulations, both in its nearly symmetric shape and its scale. We note that, while the standard deviation of displacement was input into the TKdV theory, the slope standard deviation was not. Downstream, both the experimental and theoretical slope distributions  spread significantly. Interestingly, in the experiments, it is not the standard deviation that increases significantly upon crossing the depth change ($\std(\eta_x)$ increases from 0.15 to 0.17), but instead the excess kurtosis, which jumps from a value of 0.46 upstream to 3.4 downstream. Likewise, in the TKdV theory, $\std(\eta_x)$ grows from 0.13 upstream to 0.33 downstream, and $\kurt(\eta_x)$ grows from -0.04 to 0.50. Thus, the theoretically predicted jump in kurtosis, while not as extreme as that measured in experiments, is quite significant. These elevated levels of kurtosis are associated with the distinct appearance of the distributions, most notably the flat tails that appear in both the experimental and theoretical distributions, and also the pointy center that is most pronounced in the experiments.

	Distinct differences in the shapes of the downstream distributions, however, are also visible. While the theoretical distributions remain symmetric, the experimental distributions skew towards negative slope. As mentioned earlier, this skewness is consistent with a right-moving wave of steep leading surface (i.e.~negative slope), which is a phenomenon apparently not captured by the TKdV simulations.
Finally, it should be commented that the experimental measurements of slope involve numerical differentiation of the free surface extracted from optical images, which naturally introduces error. Furthermore, high-order statistics, such as kurtosis, can be very sensitive to noise and measurement error. Hence, the experimental statistics of slope, while highly valuable, should be regarded with some level of caution.

%----------------------------------------------------------------------------%
% Wave dynamics
%----------------------------------------------------------------------------%
\subsection{Wave dynamics and analysis of time scales}

@@

These observations can be rationalized with some simple analysis of the structure of TKdV. Upstream of the ADC, since $\drat = 1$, the nonlinear effects are relatively weak and the dynamics are dominated by dispersion. Linearizing \eqref{TKdV} and introducing the ansatz $u_k = e^{i(kx - \omega_k t)}$ produces the dispersion relation for angular frequency $\omega_k = -C_2 k^3$, or for phase velocity $c_k = \omega_k/k = -C_2 k^2$ \cite{majdaqi2019}. In particular, the phase velocity is strictly negative, implying that waves can only propagate leftwards, as is consistent with Fig.~\ref{trajectories}(a).
% Further details, see page 6 of Majda Qi, J Stat Phys.
%$C_3 = 0.20$ and $C_2 = 0.024$.

@@

	We remind the reader that the KdV framework tracks the long-time evolution of waves in a reference frame moving with the characteristic speed $c = \sqrt{g \depth}$ from linear theory. Hence, the above discussion of left-or-right going waves cannot be interpreted in the context of the experiments without an appropriate Galilean transformation. Hence, in the laboratory frame, all waves indeed propagate unidirectionally, from left to right, with a speed near $c$. The directions and speeds calculated by the TKdV framework simply quantify the deviation of the true wave speed from $c$.

@@

	These estimates raise an important question: if a distance of 22 meters is required for significant wave evolution under the KdV framework, how do the experiments exhibit substantial changes in wave statistics a much shorter distance downstream of the ADC? The results shown in Fig.~\ref{skew-kurt} help resolve this question. This figure shows the evolution of surface-displacement skewness (left) and kurtosis (right) computed by the same deterministic TKdV simulations as pictured in Fig.~\ref{trajectories}. The bold black curve shows the ensemble mean (ensemble size of 1000), while the faint gray curves show the skewness and kurtosis of 100 individual trajectories to give a sense for the variation involved. Importantly, the skewness and kurtosis evolve on a much shorter timescale than the aforementioned $t=7$. More precisely, skewness and kurtosis have already saturated to their asymptotic value by $t = 1$, and reach half of that value by $t = 0.18$. These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order as the relevant distances in the experiments. Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter its statistical distributions. The timescale for this reconfiguration is much shorter than the time required for waves to cross the entire periodic domain.

 %----------------------------------------------------------------------------%
% Skewness Ratio 
%----------------------------------------------------------------------------%
\subsection{Explicit formula for outgoing skewness and experimental confirmation}

We now discuss arguably the most novel single result of the manuscript. In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}. More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$. Here, we recap the derivation of this formula and then test the prediction against new, direct experimental measurements.

@@

	Our next task is to convert this formula to physical variables so that it can be tested against experimental data. First, following the definition of $\Hthree$ in \eqref{H3H2} and the definition, $u = \eta/(\pi^{1/2} \etastd)$, a straightforward calculation gives
\begin{align}
\label{H3convert}
\meandn{\Hthree} = \frac{\pi}{3} \meandn{u^3} = 
\frac{1}{3 \pi^{1/2}} \frac{\meandn{\eta^3}}{ \etastd^3} = 
\frac{1}{3 \pi^{1/2}} \skwdn(\eta)
\end{align}
This equation expresses a direct relationship between the expected value of $\Hthree$ and the outgoing skewness of surface displacement. Next, we must convert the $\Htwo$-quantities, which involve surface slope. 
Using the definition of $\Htwo$ in \eqref{H3H2} and $\lengthscale_+$ in \eqref{scales} gives
\begin{align}
\label{H2convert}
&\meanupdn{\Htwo} = \pi \meanupdn{ u_x^2} = 
\left( \frac{\lamfac \lamdn}{2 \pi \etastd} \right)^2 \varupdn(\eta_x)
\end{align}
We note that the downstream scale $\lengthscale_+$ is used in this conversion, since it is the downstream Hamiltonian that is matched in \eqref{statmatch}. This formula links the expected value of $\Htwo$ to the variance of surface slope.

@@

	Fortunately, it is a relationship that can be tested directly against the experimental measurements, specifically the type of data that was presented in Fig.~\ref{ExpSpatialStats}. Formula \eqref{SkewPred} draws focus to the spatial variation of displacement skewness, which we replot for convenience in Fig.~\ref{SkewRat}(a), and the variance of surface slope, shown in Fig.~\ref{SkewRat}(b). To test \eqref{SkewPred}, we must select a representative downstream position to evaluate $\skwdn(\eta)$ and $\vardn(\eta_x)$, and a representative upstream position for $\varup(\eta_x)$. In short, we select the same representative locations  used to evaluate the histograms in Figs.~\ref{ExpDiagStats}, \ref{uhist}, and \ref{slopehist}, namely $x = 15$ cm downstream and $x = -9$ cm upstream. These positions are indicated by the blue and green vertical dashed lines in Fig.~\ref{SkewRat}(a)--(b). Recall that $x = 15$ cm corresponds to the peak of the downstream skewness and kurtosis. Meanwhile, we see in Fig.~\ref{SkewRat}(b) that $x = -9$ cm corresponds to a slight dip in $\var(\eta_x)$.

	With these locations selected, we evaluate the ratio $\skw(\eta) / \Delta \var(\eta_x)$ for each experiment and plot the result against dimensionless wave amplitude, $\epsup = \etastd / \dup$ on a log-log scale in Fig.~\ref{SkewRat}(c). While the data shows significant variation at small amplitudes, it shows a well defined decreasing trend at larger amplitudes. In particular, the red dashed line shows the power-law $\epsup^{-3}$ predicted by \eqref{SkewPred}, which captures the experimental trend remarkably well.

@@

	This manuscript extends the parallel experimental and modeling efforts of Bolles et al.~\cite{bolles2019} and Majda et al.~\cite{majda2019} concerning the emergence of anomalous wave statistics from abrupt changes in bottom topography. The theoretical framework is based on deterministic and statistical analysis of the TKdV system, with exploitation of the Hamiltonian structure and the associated invariant Gibbs measures. The theory depends crucially on matching the incoming and outgoing invariant measures at the abrupt depth change, so as to link the outgoing dynamics to the incoming ones. Throughout, we have emphasized the synergy between the experiments and theory, with experimental data informing theoretical advancements and model predictions motivating new experimental measurements.

	Careful calibration of the inverse temperature against the experimental data allowed for a detailed comparison between experimental and theoretically predicted wave statistics. The outgoing distributions of surface displacement predicted by the TKdV framework capture in remarkable detail the experimental measurements. We have extended this statistical analysis to surface slope, a line of inquiry motivated by the importance of $\Htwo$, the slope standard deviation, in the theory. The comparison of slope statistics shows intriguing similarities, while also some quantitative differences in the shape and skewness of the outgoing distributions.

	Finally, Majda et al.~\cite{majda2019} derived an explicit formula for the skewness of the outgoing wave field, which we have tested against direct experimental measurements. Specifically, the formula predicts how the ratio of displacement skewness to slope variance, $\skw(\eta) / \Delta \var(\eta_x)$, depends on wave amplitude. The experimental measurements conclusively confirm the inverse-cube dependence on wave amplitude, thus highlighting the predictive power of the TKdV statistical mechanics framework.

	We note that since the ratio, $\skw(\eta) / \Delta \var(\eta_x)$, exhibits such a clear dependence on input parameters, it could prove a useful diagnostic for anomalous wave observations in the ocean. In particular, this quantity seems to provide a signature for anomalous waves that are triggered by abrupt depth changes. Examination of this ratio in field data is an exciting avenue for future research.
